\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2019-02-15 archive.tex
changed color scheme in cycle figure
better figures for learning curves

An age-old dream within AI is a machine that learns and reasons by
writing its own programs.  This vision stretches back to the
1960's~\cite{solomonoff1964formal} and and is central to much of
what it would take to build machines that learn and think like humans.
Computational models of cognition often explain the flexibility and
richness of human thinking in terms of program learning: from everyday
thinking and problem solving (motor program induction as an account of
recognition and generation of handwriting and
speech~\cite{lake2015human}; functional programs as a model of natural
language semantics~\cite{SOMETHING}) to learning problems that unfold
over longer developmental time scales: the child's acquisition of
intuitive theories (of kinship, taxonomy, etc.)~\cite{Ullman2012} and
natural language grammar~\cite{DBLP:journals/cogsr/SchmidK11}, to name
just a few.  An outstanding challenge, however, is to engineer
program-learners that display the same level of domain-generality as
the humans they are meant to model.

@@

Each wake/sleep cycle creates new DSL components that build on
components learned in earlier sleep cycles, growing a DSL with nested 
hierarchies of code.  We identify this cumulative nesting of
abstractions as a kind of deep representation
learning~\cite{lecun2015deep}.  Figure~\ref{exampleDSL} diagrams a
subset of these learned networks (the DSL). For example, the system
learns to sort sequences of numbers by invoking a DSL component 4
layers deep, or draws the leftmost pairs of images in
Figure~\ref{exampleDSL} using a depth-3 component.  For this reason we
refer to \system as an instance of `deep program learning'.

Because any learning problem can in principle be cast as program
induction, it is important to delimit our focus.  In contrast to
computer assisted programming~\cite{solar2008program} or genetic
programming~\cite{DBLP:books/daglib/0070933}, our goal is not to
automate software engineering, or to synthesize large bodies of code
starting from scratch.  Ours is a basic AI goal: capturing the human
ability to learn to think flexibly and efficiently in new domains ---
to learn what you need to know about a domain so you don't have to
solve new problems starting from scratch.  We are focused on problems
that people solve relatively quickly, once they acquire the relevant
domain expertise.  These correspond to tasks solved by short programs
--- if you have an expressive DSL.

@@

\system~takes as its goal to acquire domain-specific expertise, which
it learns by solving a collection of programming \textbf{tasks}.  It
alternatingly finds programs that solve tasks (Wake --
Figure~\ref{threeCycles} top); improves its DSL by discovering and
reusing domain-specific subroutines (Consolidation --
Figure~\ref{threeCycles} left); and trains a neural network that
efficiently guides search for programs in the DSL (Dreaming --
Figure~\ref{threeCycles} right).  The learned DSL acts as a a prior on
programs likely to solve tasks in the domain, while the neural net
looks at a specific task and produces a ``posterior'' for programs
likely to solve that specific task (Figure~\ref{threeCycles} middle).
The neural network thus functions as a \textbf{recognition model}
supporting a form of approximate Bayesian program induction, jointly
trained with a \textbf{generative model} for programs encoded in the
DSL, in the spirit of the Helmholtz machine~\cite{hinton1995wake}. The
recognition model ensures that searching for programs remains
tractable even as the DSL (and hence the search space for programs)
expands.  The generative model, or DSL, distills out common
abstractions across programs discovered during waking, growing a
network of increasingly deep and specialized domain-specific concepts
(Figure~\ref{exampleDSL}, bottom rows).

We view these wake sleep/cycles as an approximate inference algorithm that starts with a collection of tasks, written $X$, and infers both a program for each task, as well
as a distribution over programs encoded by a DSL, written
$\mathcal{D}$.  We equip $\mathcal{D}$ with a learned weight vector
$\theta$, and together $(\mathcal{D},\theta)$ define a generative
model over programs (Appendix~\ref{sampleProgram}).  We perform
maximum a posteriori (MAP) inference of the DSL and programs given
$X$.  Writing $J$ for the joint probability of $(\mathcal{D},\theta)$
and $X$, we seek the generative model $\mathcal{D}^*,\theta^*$ solving:
\begin{align}\label{intractableObjectives}
\nonumber  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]&\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta &\qquad
  \theta^* =\argmax_\theta J(\mathcal{D}^*,\theta)
\end{align}
%% where $\probability[x|p]$ scores the likelihood of a task
%% $x\in X$ given a program $p$.\footnote{For example, for list
%%   processing, the likelihood is 1 if the program predicts the observed
%%   outputs on the observed inputs, and 0 otherwise; when learning a
%%   generative model or probabilistic program, the likelihood is the
%%   probability of the program sampling the observation.}
Evaluating Eq.~\ref{intractableObjectives}
entails marginalizing over the infinite set of all programs -- which is impossible.
We make a particle-based approximation to Eq.~\ref{intractableObjectives}
and instead marginalize over a finite \textbf{beam} of programs,
with one beam per task, collectively written $\left\{\mathcal{B}_x \right\}_{x\in X}$.
This particle-based approximation is written $\lowerBound (\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})$
and acts as a lower bound on the joint density:
\begin{align}
 J(\mathcal{D},\theta)\geq \lowerBound  (\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})\triangleq\probability[\mathcal{D},\theta]\prod_{x\in X} \sum_{p\in \mathcal{B}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]\text{, where $|\mathcal{B}_x|$ is small}
\end{align}
Wake and sleep cycles correspond to alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{B}_x \right\}_{x\in X}$  (\textbf{Wake})
and $(\mathcal{D},\theta)$ (\textbf{Consolidation}):
\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the beams.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the beams so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]\propto\probability[p|x,\mathcal{D},\theta]$ 
is large, i.e., programs with high posterior probability,
or programs which are a priori likely and which solve the task.
We find programs solving a task by
enumerating programs
from the DSL in decreasing order of their probability
under the recognition model,
and then check if a program $p$ assigns positive probability to a task
($\probability[x|p] > 0$); if so, we incorporate $p$ into the beam
$\mathcal{B}_x$. We represent programs as polymorphicly typed $\lambda$-calculus expressions,
an expressive formalism
including conditionals,
variables,
higher-order recursive functions,
and the ability to define new functions.
\\\noindent \textbf{Sleep (Consolidation): Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{B}_x \right\}_{x\in X}$ is held fixed and the problem is to search the discrete space of DSLs and find one maximizing $\int \lowerBound \;\mathrm{d}\theta$,
and then update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})$.

This 3-phase inference procedure works through two distinct kinds of bootstrapping.
During each sleep cycle the next DSL
bootstraps off the primitives learned
during earlier cycles,
growing an increasingly deep learned program representation (`deep program learning').
In tandem the DSL and recognition model
bootstrap each other:
a more finely tuned
DSL yields
richer dreams for the recognition model to learn from,
while a more accurate recognition model
solves more tasks during waking which then feed into the next DSL.
\begin{figure}
  \begin{tikzpicture}[line width=0.4mm]
    \draw[fill=white!5!white] (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
    \node at (5.5,1.25) {\textsc{\textbf{Wake}}};

@@

    \node at (0,-4.75) {\textbf{\textsc{Sleep: Consolidation}}};
    \draw[fill=white!5!white] (-3,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-3,-13) -- (-3,-5);
    \node at (12,-4.75) {\textbf{\textsc{Sleep: Dreaming}}};
    \draw[fill=white!5!white] (15,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (15,-13) -- (15,-5);

@@

Mathematically this compression takes the form of
finding the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{overviewSection}).
We replace this marginal with an AIC approximation~\cite{akaike1998information}
and marginalize over refactorings of programs in the beams, minimizing the following expression,
which can be interpreted as a kind of compression:
\begin{equation}
\underbrace{-\log \probability[\mathcal{D}] + \min_{\theta}\Bigg(-\log
\probability[\theta|\mathcal{D}] + \|\theta\|_0}_{\text{Description length of }(\mathcal{D},\theta)} +\sum_{x\in
  X}\underbrace{-\log \sum_{\substack{p:\\
      \exists p'\in
  \mathcal{B}_x: p\text{ refactors into }p'}}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
\label{AIC}
\end{equation}
But a program has infinitely many possible refactorings, rendering
Eq.~\ref{AIC} intractable.  Rather than consider every refactoring we
bound the number of $\lambda$-calculus evaluation steps separating a
program from its refactoring.  Now the number of refactorings is
finite but astronomically large: Figure~\ref{mapFactor}A diagrams a
problem where the agent rediscovers the higher-order function
\code{map} starting from the basics of Lisp and the Y-combinator, but
where there are approximately $10^{14}$ possible refactorings -- a quantity
that grows exponentially both as a function of program size and a
function of the bound on evaluation steps. How can we tame this
combinatorial explosion?

A structure called a \textbf{version space
  algebra}~\cite{lau2001programming,mitchell1977version,polozov2015flashmeta}
neatly resolves this exponential growth.  A version space is a
tree-shaped data structure that compactly represents a large set of
programs and supports efficient set operations like union,
intersection, and membership checking.  In
Appendix~\ref{appendixVersion}, we give a dynamic program that takes
as input a program and then outputs a version space containing its
refactorings. Figure~\ref{mapFactor}B diagrams a subtree of a
version space containing refactorings of a small program. Our
technique is substantially more efficient than explicitly representing
the space of possible refactorings: for the example in
Figure~\ref{mapFactor}A, we represent the space of refactorings using
a version space with $10^6$ nodes, which encodes $10^{14}$
refactorings. Appendix~\ref{appendixCompression} specifies how we combine this probabilistic and symbolic machinery to update the DSL.
At a high level, our approach is to search locally through the space 
of DSLs, proposing small changes until Eq.~\ref{AIC}
fails to decrease.

@@

\noindent\textbf{Symmetry breaking.} A good DSL not only exposes high-level building blocks, but also
carefully restricts the ways in which those building blocks are
allowed to compose.  For example,
%% a DSL for list manipulation should
%% contain both the empty list and a routine for appending lists, but
%% should not allow appending the empty list.  Similarly 
a DSL for arithmetic should disallow adding zero, or force
right-associative addition.  These restrictions break symmetries in
the space of programs.  A bigram parameterization of the recognition
model, combined with the $\mathcal{L}^{\text{MAP}}$ training
objective, interact in a way that breaks symmetries in the program
space, allowing the agent to more efficiently explore the space of
programs.  This interaction occurs because the bigram parameterization
can disallow DSL primitives depending on their local syntactic
context, while the $\mathcal{L}^{\text{MAP}}$ objective forces all
probability mass onto a single member of a set of syntactically
distinct but semantically equivalent expressions
(Appendix~\ref{recognitionAppendix}).
We experimentally confirm this symmetry-breaking
behavior by training recognition models that minimize either
$\mathcal{L}^{\text{MAP}}$/$\mathcal{L}^\text{posterior}$ and which
use either a bigram parameterization/unigram\footnote{In the unigram variant $Q$ predicts a $|\mathcal{D}| + 1$-dimensional vector: $Q(p|x) = \probability[p|\mathcal{D},\theta_i = Q_i(x)]$,
  and was used in our prior work~\cite{ecc}} parameterization.
Figure~\ref{symmetry} (right) shows the result of training $Q$ in these four regimes
and then sampling programs.
On this particular run,
the combination of
bigrams and $\mathcal{L}^{\text{MAP}}$ learns to
avoid adding zero and associate addition to the right ---
different random initializations
lead to either right or left association.

@@

\subsubsection{List Processing}\label{listSection}
We took 218 list manipulation tasks from our previous work~\cite{ecc},
each with 15 input/output examples.  In solving these tasks, the
system composed 16 new DSL components, and discovered multiple
higher-order functions. Each round of memory consolidation built on
components discovered in earlier sleep cycles --- for example the
agent first learns the higher-order function \code{filter}, uses
\code{filter} to learn to take the maximum element of a list, then
uses that routine to learn a new component for extracting the
$n^{\text{th}}$ largest element of a list, which it finally uses to
learn how to sort a list of numbers (Figure~\ref{initialExampleDSL}).
This incremental, modular learning of deep hierarchies of DSL
components occurs because of the alternation between code writing
(during waking) and code refactoring (during the consolidation phase
of sleep).

@@

Inside its learned DSL we find interpretable parametric drawing
routines corresponding to the families of visual objects in its
training data, like polygons, circles, and spirals
(Figure~\ref{logoPrimitives}, left) -- without supervision the agent
has learned the basic types of objects in its visual world. It
additionally learns more abstract visual relationships, like
rotational symmetry, which it models by incorporating a higher-order
function into its DSL (Figure~\ref{logoPrimitives}, right).  This
abstraction comes from jointly compressing programs for all of the
training images into its DSL.

@@

\begin{figure}\centering
  \includegraphics[width = 0.3\textwidth]{figures/logo16.png}
  \includegraphics[width = 0.3\textwidth]{figures/tower9.png}
  \includegraphics[width = 0.3\textwidth]{figures/sr16.png}
  \caption{Three domains where the agent infers a program from visual input. \textbf{Left}: 16 (out of 145) LOGO graphics tasks. Agent writes a program controlling a `pen' that draws the target picture. \textbf{Middle}: 9 (out of 112) tower building tasks. Agent writes a program controlling a `hand' that builds the target tower. \textbf{Right}: 16 (out of 200) symbolic regression tasks. Agent writes a program containing continuous real numbers that fits the points along the curve.}\label{visualSpecs}
\end{figure}
\begin{figure}
\centering  \begin{tabular}{cc}
    \toprule
    Parametric drawing routines&Higher-order drawing routine\\\midrule
    \begin{tabular}{rl}
      Semicircle:& \raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_24.png}}\\
      Circles:
      &\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_23.png}}\\
      Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_25.png}}\\
      Greek Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_10.png}}\\
      S-Curves:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_11.png}}        \\
        Polygons \& Stars:
&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_12.png}}
    \end{tabular}
    &
    \begin{tabular}{c}
      Rotational Symmetry:\\
          \includegraphics[width = 0.35\textwidth]{figures/rotationalmontage.png}
      \end{tabular}

@@

\begin{figure}
  \begin{tabular}{ccc}
    \includegraphics[width = 5cm]{figures/logoLearningCurve_batch.png}&
    \includegraphics[width = 5cm]{figures/textLearningCurve_batch.png}&
    \includegraphics[width = 5cm]{figures/listLearningCurve_batch.png}\\\\
    \includegraphics[width = 5cm]{figures/towerLearningCurve_batch.png}&
    \includegraphics[width = 5cm]{figures/rationalLearningCurve_batch.png}&
    \includegraphics[width = 5cm]{figures/regexLearningCurve_batch.png}
  \end{tabular}
  \caption{Test set performance across wake/sleep iterations. Orange: \systemEnding. Teal: No recognition model. Purple: No DSL learning. Solid lines: \% tasks solved within 10 minute timeout. Dashed lines: average search time. Shaded region is $\pm 1$ stddev search time.}\label{learningCurves}
\end{figure}

@@

>>>>

An old dream within AI is a machine that learns and reasons by
writing its own programs.  This vision stretches back to the
1960's~\cite{solomonoff1964formal} and and is central to much of
what it would take to build machines that learn and think like humans.
Computational models of cognition often explain the flexibility and
richness of human thinking in terms of program learning: from everyday
thinking and problem solving (motor program induction as an account of
recognition and generation of handwriting and
speech~\cite{lake2015human}; functional programs as a model of natural
language semantics~\cite{SOMETHING}) to learning problems that unfold
over longer developmental time scales: the child's acquisition of
intuitive theories (of kinship, taxonomy, etc.)~\cite{Ullman2012} and
natural language grammar~\cite{DBLP:journals/cogsr/SchmidK11}, to name
just a few.  An outstanding challenge, however, is to engineer
program-learners that display the same level of domain-generality as
the humans they are meant to model.

@@

Each wake/sleep cycle creates new DSL components that build on
components learned in earlier sleep cycles, growing a DSL with nested 
hierarchies of code.  We identify this cumulative nesting of
abstractions as a variety of deep representation
learning~\cite{lecun2015deep}.  Figure~\ref{exampleDSL} diagrams a
subset of these learned networks (the DSL). For example, the system
learns to sort sequences of numbers by invoking a DSL component 4
layers deep, or draws the leftmost pairs of images in
Figure~\ref{exampleDSL} using a depth-3 component.  For this reason we
refer to \system as an instance of `deep program learning'.

Our goal with \system is to engineer a system that develops domain
expertise in humanlike ways. This involves learning both declarative and
procedural knowledge, like a DSL and a recognition model, but includes
other features of the development of expertise:
humans can become experts in many fields,
and so we evaluate our algorithm across six different problem domains;
a human expert doesn't become an expert overnight,
and needs more than a handful of example problems to learn from,
but doesn't need millions of examples;
similarly our algorithm's
learning trajectory unfolds over a series of wake/sleep cycles
requiring around a hundred problems per domain.

@@

\system~takes as its goal to acquire domain-specific expertise, which
it learns by solving a collection of programming \textbf{tasks}.  It
alternatingly finds programs that solve tasks (Wake --
Figure~\ref{threeCycles} top); improves its DSL by analyzing programs
found during waking (Consolidation -- Figure~\ref{threeCycles} left);
and trains a neural network that efficiently guides search for
programs in the DSL (Dreaming -- Figure~\ref{threeCycles} right).  The
learned DSL acts as a a prior on programs likely to solve tasks in the
domain, while the neural net looks at a specific task and produces a
``posterior'' for programs likely to solve that specific task
(Figure~\ref{threeCycles} middle).  The neural network thus functions
as a \textbf{recognition model} supporting a form of approximate
Bayesian program induction, jointly trained with a \textbf{generative
  model} for programs encoded in the DSL, in the spirit of the
Helmholtz machine~\cite{hinton1995wake}. The recognition model ensures
that searching for programs remains tractable even as the DSL (and
hence the search space for programs) expands.  The generative model,
or DSL, distills out common abstractions across programs discovered
during waking, growing a network of increasingly deep and specialized
domain-specific concepts (Figure~\ref{exampleDSL}, bottom rows).

%% Formally, $EC^2$ takes as input a collection of problems, written $X$, which can be solved by programs, and infers a domain specific language -- written $\mathcal{D}$ -- which functions as a generative model over programs likely to solve problems in the domain, and which we identify as an expert's a priori beliefs about what good solutions look like. Alongside this generative model it trains a neural recognition model, in the spirit of the Helmholtz machine or wake-sleep algorithm~\cite{dayan1995helmholtz}. The recognition model takes as input a specific problem and quickly predicts an approximate posterior over programs likely to solve that problem. Writing $p_x$ for the program solving problem $x\in X$ and $Q(p|x)$ for the distribution predicted by the recognition model, $EC^2$ iteratively (and approximately) solves for
%% \begin{align}
%% p_x& = \argmax_p \probability[x|p]\probability[p|\mathcal{D}]\label{wakingObjective}\\
%% \mathcal{D}& = \argmax_\mathcal{D}\probability\left[\mathcal{D} \right]\prod_{x\in X}\sum_{\substack{p:\\p\text{ found during Explore}}} \probability[p|\mathcal{D}]\label{sleepg}\\
%% Q(p|x)&\approx \probability[p|x,\mathcal{D}]\label{dreamingObjective}
%% \end{align}

These wake sleep/cycles function as an approximate inference algorithm
that observes a collection of tasks, written $X$, and infers both a
program solving each task, as well as a distribution over programs
encoded by a DSL, written $\mathcal{D}$.  We equip $\mathcal{D}$ with
a learned weight vector $\theta$, and together $(\mathcal{D},\theta)$
define a generative model over programs
(Appendix~\ref{sampleProgram}).  Writing $Q(p|x)$ for the approximate
posterior predicted by the recognition model, we iteratively (and
approximately) solve for

\begin{array}{lr}
  \begin{array}{l}  p_x = \argmax_p \probability[x|p]\probability[p|\mathcal{D}^*,\theta^*]\end{array}
    &\text{\emph{Wake}}\\\\
\begin{array}{rl}  
\mathcal{D}^* = \argmax_\mathcal{D}&\int\probability\left[\mathcal{D},\phantom{^*}\theta \right]\prod_{x\in X}\sum_{\substack{p\text{ found during waking}}} \probability[x|p]\probability[p|\mathcal{D},\phantom{^*}\theta]\;\text{d}\theta\\
\theta^* = \argmax_\theta&\phantom{\int}\probability\left[\mathcal{D}^*,\theta \right]\prod_{x\in X}\sum_{p\text{ found during waking}} \probability[x|p]\probability[p|\mathcal{D}^*,\theta]
 \end{array}
&\text{\emph{Consolidation sleep}}\\\\
\begin{array}{l} Q(p|x)\approx \probability[p|x,\mathcal{D}^*,\theta^*]\end{array}
  &\text{\emph{Dream sleep}}
\end{array}

\noindent which serves to maximize a lower bound on the posterior over $(\mathcal{D},\theta)$ given $X$ (Appendix~\ref{probabilisticAppendix}).

This 3-phase inference procedure works through two distinct kinds of bootstrapping.
During each sleep cycle the next DSL
bootstraps off the primitives learned
during earlier cycles,
growing an increasingly deep learned program representation.
In tandem the DSL and recognition model
bootstrap each other:
a more finely tuned
DSL yields
richer dreams for the recognition model to learn from,
while a more accurate recognition model
solves more tasks during waking which then feed into the next DSL.

Waking consists of searching for task-specific programs with high
posterior probability, or programs which are a priori likely and which
solve a task.  We find programs solving a task by enumerating
programs from the DSL in decreasing order of their probability under
the recognition model, and then check if a program $p$ assigns
positive probability to a task ($\probability[x|p] > 0$). We represent
programs as polymorphicly typed $\lambda$-calculus expressions, an
expressive formalism including conditionals, variables, higher-order
recursive functions, and the ability to define new functions.

\begin{figure}
  \begin{tikzpicture}[line width=0.4mm]
    \draw[fill=teal!5!white] (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
    \node at (5.5,1.25) {\textsc{\textbf{Wake}}};

@@

    \node at (0,-4.75) {\textbf{\textsc{Sleep: Consolidation}}};
    \draw[fill=teal!5!white] (-3,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-3,-13) -- (-3,-5);
    \node at (12,-4.75) {\textbf{\textsc{Sleep: Dreaming}}};
    \draw[fill=teal!5!white] (15,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (15,-13) -- (15,-5);

@@

Mathematically this compression takes the form of
finding the DSL maximizing $\int \probability[\mathcal{D},\theta]\probability[X|\mathcal{D},\theta]\;\mathrm{d}\theta$ (Sec.~\ref{overviewSection}).
We replace this marginal with an AIC approximation~\cite{akaike1998information}
and marginalize over refactorings of programs found during waking, minimizing the following expression,
which can be interpreted as a kind of compression:
\begin{equation}
\underbrace{-\log \probability[\mathcal{D}] + \min_{\substack{\theta\\\\\\\\}}\Bigg(-\log
\probability[\theta|\mathcal{D}] + \|\theta\|_0}_{\text{Description length of }(\mathcal{D},\theta)} +\sum_{x\in
  X}\underbrace{-\log \sum_{\substack{p \text{ a refactoring of } p'\\ p'\text{ found during waking}}}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
\label{AIC}
\end{equation}
But a program has infinitely many possible refactorings, rendering
Eq.~\ref{AIC} intractable.  Rather than consider every refactoring we
bound the number of $\lambda$-calculus evaluation steps separating a
program from its refactoring.  Now the number of refactorings is
finite but astronomically large: Figure~\ref{mapFactor}A diagrams a
problem where the agent rediscovers the higher-order function
\code{map} starting from the basics of Lisp and the Y-combinator, but
where there are approximately $10^{14}$ possible refactorings -- a quantity
that grows exponentially both as a function of program size and a
function of the bound on evaluation steps. How can we tame this
combinatorial explosion?

To resolve this exponential growth we introduce a new data structure
combining ideas from version space
algebras~\cite{lau2001programming,mitchell1977version,polozov2015flashmeta}
and equivalence graphs~\cite{tate2009equality}. A version space is a
tree-shaped data structure that compactly represents a large set of
programs and supports efficient set operations like union,
intersection, and membership checking, while equivalence graphs are
data structures that track semantic equivalences between program
subexpressions.  In Appendix~\ref{appendixVersion}, we give a dynamic
program that takes as input a program and then outputs a version space
containing its refactorings while tracking semantically equivalent
subexpressions. Figure~\ref{mapFactor}B diagrams a subtree of a
version space containing refactorings of a small program. Our
technique is substantially more efficient than explicitly representing
the space of possible refactorings: for the example in
Figure~\ref{mapFactor}A, we represent the space of refactorings using
a version space with $10^6$ nodes, which encodes $10^{14}$
refactorings. Appendix~\ref{appendixCompression} specifies how we
combine this probabilistic and symbolic machinery to update the DSL.
At a high level, our approach is to search locally through the space
of DSLs, proposing small changes until Eq.~\ref{AIC} fails to
decrease.

@@

\noindent\textbf{Symmetry breaking.} A good DSL not only exposes high-level building blocks, but also
carefully restricts the ways in which those building blocks are
allowed to compose.  For example,
%% a DSL for list manipulation should
%% contain both the empty list and a routine for appending lists, but
%% should not allow appending the empty list.  Similarly 
a DSL for arithmetic should disallow adding zero, or force
right-associative addition. A bigram parameterization of the recognition
model, combined with the $\mathcal{L}^{\text{MAP}}$ training
objective, interact in a way that breaks symmetries like these, allowing the agent to more efficiently explore the space of
programs.  This interaction occurs because the bigram parameterization
can disallow DSL primitives depending on their local syntactic
context, while the $\mathcal{L}^{\text{MAP}}$ objective forces all
probability mass onto a single member of a set of syntactically
distinct but semantically equivalent expressions
(Appendix~\ref{recognitionAppendix}).
We experimentally confirm this symmetry-breaking
behavior by training recognition models that minimize either
$\mathcal{L}^{\text{MAP}}$/$\mathcal{L}^\text{posterior}$ and which
use either a bigram parameterization/unigram\footnote{In the unigram variant $Q$ predicts a $|\mathcal{D}| + 1$-dimensional vector: $Q(p|x) = \probability[p|\mathcal{D},\theta_i = Q_i(x)]$,
  and was used in our prior work~\cite{ecc}} parameterization.
Figure~\ref{symmetry} (right) shows the result of training $Q$ in these four regimes
and then sampling programs.
On this particular run,
the combination of
bigrams and $\mathcal{L}^{\text{MAP}}$ learns to
avoid adding zero and associate addition to the right ---
different random initializations
lead to either right or left association.

@@

\subsubsection{List Processing}\label{listSection}
We took 218 list manipulation tasks from our previous work~\cite{ecc},
each with 15 input/output examples.  In solving these tasks, the
system composed 16 new DSL components, and discovered multiple
higher-order functions. Each round of memory consolidation built on
components discovered in earlier sleep cycles --- for example the
agent first learns the higher-order function \code{filter}, uses
\code{filter} to learn to take the maximum element of a list, then
uses that routine to learn a new component for extracting the
$n^{\text{th}}$ largest element of a list, which it finally uses to
solve a task involving sorting a list of numbers (Figure~\ref{initialExampleDSL}).
This incremental, modular learning of deep hierarchies of DSL
components occurs because of the alternation between code writing
(during waking) and code refactoring (during the consolidation phase
of sleep).

@@

Inside its learned DSL we find interpretable parametric drawing
routines corresponding to the families of visual objects in its
training data, like polygons, circles, and spirals
(Figure~\ref{logoPrimitives}, left) -- without supervision the agent
has learned the basic types of objects in its visual world. It
additionally learns more abstract visual relationships, like
rotational symmetry, which it models by incorporating a higher-order
function into its DSL (Figure~\ref{logoPrimitives}, right).  %% This
%% abstraction comes from jointly compressing programs for all of the
%% training images into its DSL.

@@

\begin{figure}\centering
  \includegraphics[width = 0.3\textwidth]{figures/logo16.png}
  \includegraphics[width = 0.3\textwidth]{figures/tower9.png}
  \includegraphics[width = 0.3\textwidth]{figures/sr16.png}
  \caption{Three domains where the agent infers a program from visual input. \textbf{Left}: 16 (out of 160) LOGO graphics tasks. Agent writes a program controlling a `pen' that draws the target picture. \textbf{Middle}: 9 (out of 112) tower building tasks. Agent writes a program controlling a `hand' that builds the target tower. \textbf{Right}: 16 (out of 200) symbolic regression tasks. Agent writes a program containing continuous real numbers that fits the points along the curve.}\label{visualSpecs}
\end{figure}
\begin{figure}
\centering  \begin{tabular}{cc}
    \toprule
    Parametric drawing routines&Higher-order drawing routine\\\midrule
    \begin{tabular}{rl}
      Semicircle:& \raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_24.png}}\\
      Circles:
      &\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_23.png}}\\
      Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_25.png}}\\
      Greek Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_10.png}}\\
      S-Curves:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_11.png}}        \\
        Polygons \& Stars:
&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_12.png}}
    \end{tabular}
    &
    \begin{tabular}{c}
      Rotational Symmetry:\\
          \includegraphics[width = 0.35\textwidth]{figures/rotationalmontage.png}
      \end{tabular}

@@

\begin{figure}
  \begin{tabular}{ccc}
    \begin{tabular}{c}
      \includegraphics[width = 5cm]{figures/learningCurves/text_hits.png}\\
      \includegraphics[width = 5cm]{figures/learningCurves/text_time.png}
    \end{tabular}&

    \begin{tabular}{c}
      \includegraphics[width = 5cm]{figures/learningCurves/logo_hits.png}\\
      \includegraphics[width = 5cm]{figures/learningCurves/logo_time.png}
    \end{tabular}&

    \begin{tabular}{c}
      \includegraphics[width = 5cm]{figures/learningCurves/rational_hits.png}\\
      \includegraphics[width = 5cm]{figures/learningCurves/rational_time.png}
    \end{tabular}\\\\\\

    \begin{tabular}{c}
      \includegraphics[width = 5cm]{figures/learningCurves/list_hits.png}\\
      \includegraphics[width = 5cm]{figures/learningCurves/list_time.png}
    \end{tabular}&

    \begin{tabular}{c}
      \includegraphics[width = 5cm]{figures/learningCurves/tower_hits.png}\\
      \includegraphics[width = 5cm]{figures/learningCurves/tower_time.png}
    \end{tabular}
  \end{tabular}
  \caption{Test set performance across wake/sleep iterations. Each curve is a run with a different random seed. Teal: Full model. Orange: Dreaming only (no DSL learning). Purple: Consolidation only (no recognition model). Search time plots show solid lines (time averaged over all tasks) and dotted lines (time averaged over solved tasks).}\label{learningCurves}      
\end{figure}

%% \begin{figure}
%%   \begin{tabular}{ccc}
%%     \includegraphics[width = 5cm]{figures/logoLearningCurve_batch.png}&
%%     \includegraphics[width = 5cm]{figures/textLearningCurve_batch.png}&
%%     \includegraphics[width = 5cm]{figures/listLearningCurve_batch.png}\\\\
%%     \includegraphics[width = 5cm]{figures/towerLearningCurve_batch.png}&
%%     \includegraphics[width = 5cm]{figures/rationalLearningCurve_batch.png}&
%%     \includegraphics[width = 5cm]{figures/regexLearningCurve_batch.png}
%%   \end{tabular}
%%   \caption{Test set performance across wake/sleep iterations. Orange: \systemEnding. Teal: No recognition model. Purple: No DSL learning. Solid lines: \% tasks solved within 10 minute timeout. Dashed lines: average search time. Shaded region is $\pm 1$ stddev search time.}\label{learningCurves}
%% \end{figure}

@@

\subsection{Probabilistic Formulation of \system}\label{probabilisticAppendix}
Our objective is to infer the maximum a posteriori DSL $\mathcal{D}$ and
parameters $\theta$. Writing $J$ for the joint probability of $(\mathcal{D},\theta)$,
this corresponds to solving
\begin{align}\label{intractableObjectives}
\nonumber  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]&\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta& \qquad
  \theta^* =\argmax_\theta J(\mathcal{D}^*,\theta)
\end{align}
where $\probability[x|p]$ scores the likelihood of a task
$x\in X$ given a program $p$.\footnote{For example, for list
  processing, the likelihood is 1 if the program predicts the observed
  outputs on the observed inputs, and 0 otherwise; when learning a
  generative model or probabilistic program, the likelihood is the
  probability of the program sampling the observation.}
